---
title: "Trait Analysis Simulation Example"
output: 
  html_notebook: 
    toc: yes
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


In this document, we show several trait modeling analyses using data collected in the Baviaanksloof region of the Cape Floristic Region in South Africa.

```{r, message=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(GGally)
library(ade4)
library(vegan)
library(reshape2)
library(FD)
```

Read in the simulated trait data sets.
These data were simulated using the script `construct-sim-data.Rmd`.

Following the notation of Dray and colleagues (REFS), we will have a **Q** matrix.
The **Q** matrix is the $p \times s$ table describing $s$ traits for $p$ species. 
By convention, each species is assigned a single value for each trait, usually an average if the trait is measured on a continuous scale.
In our datasets, we have multiple measurements per species, so we must *reduce* our data, resulting in a loss of some of our collected information.
Essentially, this process ignores *intra-specific* variation.
It remains unclear what role *intra-specific* variation plays in determining the patterns of trait by environment relationships in real communities.

```{r}
indXtraits <- read.csv("individualsXtraits_sim.csv")
q_sim <- indXtraits %>% group_by(spec) %>% 
  summarise(lma = mean(lma),
            c_to_n = mean(c_to_n),
            succulence = mean(succulence))
```


# Data visualization

First look at data in a non-aggregated way. 
Each data point represents measurements taken from samples collected from individual plants.

## Trait-by-trait plots

Using the `GGally::ggpairs` function.

```{r}
ggpairs(indXtraits[3:5])
```


# PCA of individual-level traits

In order to reduce dimensionality of the trait data set, we can apply principal component analysis. 

```{r}
indXtraits_pca <- rda(indXtraits[3:5], scale = TRUE)
```

Plot biplot and examine summary.

```{r}
indXtraits_pca
```

```{r, eval=FALSE}
summary(indXtraits_pca, scaling = 2)
summary(indXtraits_pca, scaling = 1)
```


```{r}
biplot(indXtraits_pca, scaling = 2, main = "PCA - scaling = 2", type = c("text", "points"))
biplot(indXtraits_pca, scaling = 2, choices = c(1,3), main = "PCA - scaling = 2", type = c("text", "points"))
```

```{r}
biplot(indXtraits_pca, scaling = 1, main = "PCA - scaling = 1", type = c("text", "points"))
```


# Species-level trait values

As outlined in the main text of the chapter, there are many challenges and problems presented when *intraspecific* trait variation is either ignored or not considered.
However, most of the 'standard' trait analyses (e.g., Fourth-corner problem, RLQ, etc.) use only species-level trait values.
Furthemore, to some degree, the assumption of indpendence among samples is violated when we combine both intra- and inter-specific sources of variation, as we have done in the PCA of traits shown above. 
For these reasons, it's worth calculating species-level mean trait values, i.e. the Q matrix described above.

### Species-level trait-trait relationships

```{r}
ggpairs(select(q_sim, -spec))
```


### PCA of species-level trait data

We can now perform a PCA of the species-level trait data.
Note that since we have calculated species-level trait mean values, we are no longer confounding intra- and inter-specific trait variation in this anlaysis.
We're not going to actually do these calculations for the simulated data since the dimenions are already very low (trait space = 3d, env space = 2d).


# Read in environmental data

The dataset of environmental conditions at the plots corresponds to the **R** matrix, following the notation of Dray and colleagues. 
This matrix is an $n \times m$ table of measurements of $m$ environmental variables taken at $n$ plots or sites.


```{r}
r_sim <- read.csv("plotXenv_sim.csv")
```

### Exploritory analysis of environmental data

Use `ggpairs` to examine distributional patterns of environmental variables and correlation among variables.

```{r}
ggpairs(r_sim[-1])
```



# Individual-level Trait by environment relationships

We can look for associations between trait values and environmental. 
This form of analysis is important in trait by environment analyses, but generally not available if species-level trait values are being used.
That is, this kind of analysis is best done for individuals.

### Make combined trait and field site environment datasets

```{r}
## Make a combined data.frame for plotting
traitsXspeciesXenv <- merge(indXtraits, r_sim, by = "plot")

## Melt this data.frame in a two-step process
traitsXspeciesXenv_m <- melt(traitsXspeciesXenv, 
                             measure.vars = c("ev1", "ev2"), 
                             variable.name = "ev_var", value.name = "ev_val")
traitsXspeciesXenv_m <- melt(traitsXspeciesXenv_m,
                             measure.vars = c("lma", "c_to_n", "succulence"), 
                             variable.name = "trait_var", value.name = "trait_val")

## Make plots
ggplot(data = traitsXspeciesXenv_m, aes(x = ev_val, y = trait_val, colour = spec)) +
  geom_point() +
  facet_grid(trait_var ~ ev_var, scales = "free") +
  geom_smooth() +
  theme_bw()
```



### Linear regression models

```{r}
traitXenv_ols_models <- 
  traitsXspeciesXenv_m %>%
  group_by(trait_var, ev_var ) %>% 
  do( mod = lm( trait_val ~ ev_val, data = . ) )

traitXenv_ols_models <- 
  mutate( traitXenv_ols_models, r_sqrd = summary(mod)$adj.r.squared )

traitXenv_ols_models_summary <-
  traitXenv_ols_models %>% 
  do(data.frame( var = names(coef(.$mod)),
                 coef(summary(.$mod))))

```



# Community Level Data

### Read in the plot-level abundance data

This is equivelant to the **L** matrix, following the notation of Dray et al..
The **L** matrix is a $n \times p$ table of abundances (or occurrences) of $p$ species in $n$ sites (or plots).

```{r}
l_sim <- read.csv("plotXspec_abund_sim.csv")
```



## RDA of CWM traits with environment


### CWM of traits

Use the `FD` package to calculate Community Weighted Mean trait values.

```{r}
## Make two data.frames, properily formated for FD functions
q_sim_mod <- q_sim[-1]
row.names(q_sim_mod) <- q_sim$spec
l_sim_mod <- l_sim[-1]
l_sim_mod <- l_sim_mod[sort(names(l_sim_mod))]


sim_dbFD_traits <-
  dbFD( x = q_sim_mod, 
        a = l_sim_mod,
        w.abun = TRUE, 
        calc.FDiv = TRUE, 
        calc.CWM = TRUE,
        corr = "cailliez" )

```


```{r}
## Get the CWM value data.frame
cwm_traits <- sim_dbFD_traits$CWM
cwm_traits$plot <- l_sim$plot
```

#### Plot CWM trait-trait relationships

```{r}
ggpairs(select(cwm_traits, -plot))
```


In the above, I calculated CWM trait values using the standard approach, where we assume that we have only one trait value for each species.
What do we get if we calculate an actual CWM using the full dataset?

```{r}
cwm_traits_direct <-
  indXtraits %>%
  group_by(plot) %>%
  summarise( lma = mean(lma),
             c_to_n = mean(c_to_n),
             succulence = mean(succulence))
```

Plot the traits from the indirect versus direct CWM measures

```{r}
plot(cwm_traits_direct$lma, cwm_traits$lma)
plot(cwm_traits_direct$c_to_n, cwm_traits$c_to_n)
plot(cwm_traits_direct$succulence, cwm_traits$succulence)
```

Plot the CWM-direct trait-trait interactions

```{r}
ggpairs(cwm_traits_direct[-1])
```



# RLQ for traits, abundance, and environment


```{r}
```



