---
title: "Trait Analysis Simulation Example"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
---

This document accompanies the book chapter **Statistical Methods for Modeling Traits**, by Matthew Aiello-Lammens and John Silander, appearing in the **Handbook of Environmental and Ecological Statistics**, edited by Gelfand, Fuentes, Hoeting, and Smith.

Here we show the trait modeling analyses discussed in the corresponding chapter, as applied to simulated data.
Details of the simulated data appear in the document `construct-sim-data.Rmd`.

```{r, message=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(GGally)
library(ade4)
library(vegan)
library(reshape2)
library(FD)
library(picante)

knitr::opts_chunk$set(fig.width=6, fig.height=6) 
```

# Read in simulated trait data

Begin by reading in the simulated individual trait data set.

```{r}
indXtraits_sim <- read.csv("individualsXtraits_sim.csv")
```

### Change names of traits 

This step is to make prettier plots.

```{r}
names(indXtraits_sim) <- c("plot", "spec", "LMA", "C_N_ratio", "Succ")
```


# Exploratory Data Analysis

Use data visualization and dimension reduction to investigate large patterns in the trait data.
In this case, each data point represents trait measurements simulated for individual plants.


## Trait-by-trait plots

Using the `GGally::ggpairs` function.

```{r}
ggpairs(indXtraits_sim[3:5])
```


## PCA of individual-level traits

In order to reduce dimensionality of the trait data set, we can apply principal component analysis. 
We carry out a PCA of the simulation data set using the `rda` function from the `vegan` package.
Note that the option `scale = TRUE` is chosen here, resulting in the data set being scaled to unit variance.

```{r}
indXtraits_sim_pca <- rda(indXtraits_sim[3:5], scale = TRUE)
```

Examine summary of PCA results

```{r}
indXtraits_sim_pca
```


```{r}
summary(indXtraits_sim_pca)
```


Examine biplots using `scaling = 1` and `scaling = 2`

### Individual-level PCA biplot - scaling = 1

```{r}
biplot(indXtraits_sim_pca, scaling = 1, main = "PCA - scaling = 1", type = c("text", "points"), col = "black")
```

### Individual-level PCA biplot - scaling = 2

```{r}
biplot(indXtraits_sim_pca, scaling = 2, main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
biplot(indXtraits_sim_pca, scaling = 2, choices = c(1,3), main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
```


# Species-level trait values

As outlined in the main text of the chapter, there are many challenges and problems presented when *intraspecific* trait variation is either ignored or not considered.
However, most of the 'standard' trait analyses (e.g., Fourth-corner problem, RLQ, etc.) use only **species-level** trait values.
For these reasons, it's worth calculating species-level mean trait values, i.e. the Q matrix described above.
The **Q** matrix is the $p \times s$ table describing $s$ traits for $p$ species. 
By convention, each species is assigned a single value for each trait, usually an average if the trait is measured on a continuous scale.
In our dataset, we have multiple measurements per species, so we must *reduce* our data, resulting in a loss of some of our collected information.

## Create a **Q** matrix

```{r}
q_sim <- indXtraits_sim %>% group_by(spec) %>% 
  summarise(LMA = mean(LMA),
            C_N_ratio = mean(C_N_ratio),
            Succ = mean(Succ))
```

### Species-level trait-trait relationships

```{r}
ggpairs(select(q_sim, -spec))
```

**NOTE:** We carried out a further PCA of the species-level traits for the CFR data set, but do not carry out a similar analysis for the simulated data set.

***

# Read in environmental data

Environmental conditions at each plot were simulated, but approximately based on the gradient of Mean Maximum Temperatuere in January ($x_1$) and rainfall seasonality ($x_2$) observed in the CFR.
See `construct-sim-date.Rmd` for more details.
The dataset of environmental conditions at the plots corresponds to the **R** matrix, an $n \times m$ table of measurements of $m$ environmental variables taken at $n$ plots or sites.


```{r}
r_sim <- read.csv("plotXenv_sim.csv")
```

### Change names for prettier plots

```{r}
names(r_sim) <- c("plot", "x1", "x2")
```

## Exploritory analysis of environmental data

Use `ggpairs` to examine distributional patterns of environmental variables and correlation among variables.

```{r}
ggpairs(r_sim[-1])
```



# Individual-level Trait by environment relationships

Because we simulated trait values for each individual in each plot, we can look for associations between trait values and environmental conditions at the individual level. 
This form of analysis is important in trait by environment analyses, but generally not available if species-level trait values are being used.

## Make combined trait and environment dataset

```{r}
## Make a combined data.frame for plotting
indXtraits_sim_env <- merge(indXtraits_sim, r_sim, by = "plot")

## Melt this data.frame in a two-step process
indXtraits_sim_env_m <- melt(indXtraits_sim_env, 
                             measure.vars = c("x1", "x2"), 
                             variable.name = "ev_var", value.name = "ev_val")
indXtraits_sim_env_m <- melt(indXtraits_sim_env_m,
                             measure.vars = c("LMA", "C_N_ratio", "Succ"), 
                             variable.name = "trait_var", value.name = "trait_val")
```

### Plot individual-level trait-by-environment relationships

Note that in the first plot we use separate line types for each species, where as in the second, we use one line for all species.

```{r}
## Make plots
ggplot(data = indXtraits_sim_env_m, aes(x = ev_val, y = trait_val, linetype = spec)) +
  geom_point() +
  facet_grid(trait_var ~ ev_var, scales = "free") +
  geom_smooth() +
  theme_bw()
```

```{r}
ggplot(data = indXtraits_sim_env_m, aes(x = ev_val, y = trait_val)) +
  geom_point() +
  facet_grid(trait_var ~ ev_var, scales = "free") +
  geom_smooth() +
  theme_bw()

```


### Linear regression models

Run one-at-a-time trait-by-environment OLS models

```{r}
traitXenv_ols_models <- 
  indXtraits_sim_env_m %>%
  group_by(trait_var, ev_var ) %>% 
  do( mod = lm( trait_val ~ ev_val, data = . ) )

traitXenv_ols_models <- 
  mutate( traitXenv_ols_models, r_sqrd = summary(mod)$adj.r.squared )

traitXenv_ols_models_summary <-
  traitXenv_ols_models %>% 
  do(data.frame( var = names(coef(.$mod)),
                 coef(summary(.$mod))))

```

```{r}
print(traitXenv_ols_models_summary)
```


## Individual-level RDA

As outlined in the chapter, we can apply RDA to the individual-level trait-environment data.

Make a dataset with standarized Y and X values

```{r}
indXtraits_sim_env_std <- indXtraits_sim_env
indXtraits_sim_env_std[3:7] <- scale(indXtraits_sim_env_std[3:7])
```

Run RDA using formula notation.

```{r}
indXtraits_sim_rda <-
  rda(formula = indXtraits_sim_env[c("LMA", "C_N_ratio", "Succ")] ~ x1 + x2, data = indXtraits_sim_env_std) 
```

Examine summary.

```{r}
summary(indXtraits_sim_rda)
```

Tri-plot - scaling = 1

```{r}
plot(indXtraits_sim_rda, type='none', scaling = 1)
points(indXtraits_sim_rda, pch=20, cex=0.8, scaling = 1)
text(indXtraits_sim_rda,display='species',col='black',axis.bp=TRUE, scaling = 1)
text(indXtraits_sim_rda,display='bp', scaling = 1)
```

Tri-plot - scaling = 2

```{r}
plot(indXtraits_sim_rda, type='none', scaling = 2 )
points(indXtraits_sim_rda, pch=20, cex=0.8, scaling = 2 )
text(indXtraits_sim_rda,display='species',col='black',axis.bp=TRUE, scaling = 2)
text(indXtraits_sim_rda,display='bp', scaling = 2)
```

Get coefficients, adjusted r^2, and perform permutation tests for this RDA

```{r}
coef(indXtraits_sim_rda)

RsquareAdj(indXtraits_sim_rda)

anova(indXtraits_sim_rda)
```


***

# Community Level Data

## Read in the plot-level abundance data

This data set is equivelant to the **L** matrix, which is an $n \times p$ table of abundances (or occurrences) of $p$ species in $n$ sites (or plots).
For the simulation data set, the abundance of each species in each plot was simulated as described in `construct-sim-data.R`.

```{r}
l_sim <- read.csv("plotXspec_abund_sim.csv")
```


## CWM trait values

Use the `FD` package to calculate Community Weighted Mean trait values.
Note that the data sets need to be manipulated to that the row names in Q match the column names in L.

```{r}
## Make two data.frames, properily formated for FD functions
q_sim_mod <- as.data.frame(q_sim[-1])
row.names(q_sim_mod) <- q_sim$spec
l_sim_mod <- l_sim[-1]
l_sim_mod <- l_sim_mod[sort(names(l_sim_mod))]


sim_dbFD_traits <-
  dbFD( x = q_sim_mod, 
        a = l_sim_mod,
        w.abun = TRUE, 
        calc.FDiv = TRUE, 
        calc.CWM = TRUE,
        corr = "cailliez" )

```


```{r}
## Get the CWM value data.frame
cwm_traits_sim <- sim_dbFD_traits$CWM
cwm_traits_sim$plot <- l_sim$plot
```

### Plot CWM trait-trait relationships

```{r}
ggpairs(select(cwm_traits_sim, -plot))
```

### Direct calculation of CWM traits for simulated data set

In the above, I calculated CWM trait values using the standard approach, where we assume that we have only one trait value for each species.
What do we get if we calculate an actual CWM using the full dataset?

```{r}
cwm_traits_sim_direct <-
  indXtraits_sim %>%
  group_by(plot) %>%
  summarise( LMA = mean(LMA),
             C_N_ratio = mean(C_N_ratio),
             Succ = mean(Succ))
```

Plot the traits from the indirect versus direct CWM measures

```{r}
## Convert trait datasets into long format
cwm_traits_sim_long <- tidyr::gather(cwm_traits_sim, "trait", "t_val", 1:3)
cwm_traits_sim_direct_long <- tidyr::gather(cwm_traits_sim_direct, "trait", "t_val_direct", 2:4)

## Join two trait data sets by plot and trait
cwm_traits_sim_comb <- inner_join(cwm_traits_sim_long, cwm_traits_sim_direct_long, by = c("plot", "trait"))

## Plot
ggplot(data = cwm_traits_sim_comb, aes(x=t_val_direct, y = t_val)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  facet_wrap(~trait, scales = "free") +
  xlab("Plot-specific CWM trait value") +
  ylab("Species-level CWM trait value") +
  theme_bw()

```

### Plot the CWM-direct trait-trait interactions

```{r}
ggpairs(cwm_traits_sim_direct[-1])
```


## RDA of CWM traits with environment (CWM-RDA)

Let's ignore the differences associated with how CWMs are calculate that are show above, and go ahead and perform RDA on the CWMs calculated using the **R** matrix.

Add the plot environmental values to the CWM matrix

```{r}
cwm_traits_sim_env <- inner_join(cwm_traits_sim, r_sim, by = "plot")
```

Setup RDA

```{r}
cwm_traits_sim_rda <-
  rda(formula = cwm_traits_sim_env[c("LMA", "C_N_ratio", "Succ")] ~ x1 + x2, data = cwm_traits_sim_env) 
```

Look at summaries and plots

```{r}
summary(cwm_traits_sim_rda)
```


Get coefficients, R^2, and run permutation tests of signficance

```{r}
coef(cwm_traits_sim_rda)

RsquareAdj(cwm_traits_sim_rda)

anova(cwm_traits_sim_rda)
```

### Tri-plot - scaling = 1

```{r}
plot( cwm_traits_sim_rda, type='none', scaling = 1)
points( cwm_traits_sim_rda, pch=20, cex=0.8, scaling = 1 )
text(cwm_traits_sim_rda,display='species',col='black',axis.bp=TRUE, scaling = 1)
text(cwm_traits_sim_rda,display='bp', scaling = 1)
```

### Tri-plot - scaling = 2

```{r}
plot( cwm_traits_sim_rda, type='none', scaling = 2 )
points( cwm_traits_sim_rda, pch=20, cex=0.8, scaling = 2 )
text(cwm_traits_sim_rda,display='species',col='black',axis.bp=TRUE, scaling = 2)
text(cwm_traits_sim_rda,display='bp', scaling = 2)
```




### Modify R, L, and Q matrices for further analyses

The R, L, and Q matrices need to be modified to that they are in the exact format needed for fourth-corner and RLQ analysis.

```{r}
r_sim_mat <- r_sim[c("x1", "x2")]
row.names(r_sim_mat) <- r_sim$plot

l_sim_mat <- select(l_sim, -plot)
row.names(l_sim_mat) <- l_sim$plot
l_sim_mat <- l_sim_mat[sort(names(l_sim_mat))]

q_sim_mat <- select(q_sim, -spec)
q_sim_mat <- as.data.frame(q_sim_mat)
row.names(q_sim_mat) <- q_sim$spec
```


## Randomization approaches to test if CWMs show signes of environmental filtering

In this example, we will use 'richness' ranomizaiton.
The work flow here is to apply the randomization/ permutation on the L matrix.
Then recalculate CWMs.
Store the CWM values for comparison to the original values later.

```{r, message=FALSE, eval = FALSE}
## Choose number of iterations
iter_n <- 999
cwm_sim_rand_all <- vector()

for (iter in 1:iter_n){
  ## Shuffle the L matrix
  l_rand <- randomizeMatrix(l_sim_mat, null.model = "richness")
  
  ## Calculate CWM values
  cwm_rand <- dbFD( x = q_sim_mat, 
                    a = l_rand,
                    w.abun = TRUE, 
                    calc.FDiv = FALSE, 
                    calc.CWM = TRUE,
                    corr = "cailliez" )
  cwm_rand <- cwm_rand$CWM
  
  ## Add a column for the iteration number
  cwm_rand$iter <- iter
  cwm_rand$plot <- row.names(cwm_rand)
  
  ## Add this iteration to the full dataset
  cwm_sim_rand_all <- rbind(cwm_sim_rand_all, cwm_rand)
}

save(cwm_sim_rand_all, file = "cwm_sim_rand_all.RData")

```


After generating random CWM sets, aggregate the data.

```{r}
## Load in the saved randomization results
load("cwm_sim_rand_all.RData")

## Calculate the mean values
cwm_rand_mean <- cwm_sim_rand_all %>%
  group_by(plot) %>%
  summarise(LMA = mean(LMA),
            C_N_ratio = mean(C_N_ratio),
            Succ = mean(Succ))
  
## Calculate the SD values
cwm_rand_sd <- cwm_sim_rand_all %>%
  group_by(plot) %>%
  summarise(LMA = sd(LMA),
            C_N_ratio = sd(C_N_ratio),
            Succ = sd(Succ))


```

Get the observed values and calculate SES

```{r}
cwm_obs_sim <- cwm_traits_sim
cwm_obs_sim <- cwm_obs_sim[order(cwm_obs_sim$plot),]
row.names(cwm_obs_sim) <- cwm_obs_sim$plot
cwm_obs_sim <- select(cwm_obs_sim, LMA:Succ)

## Check that they are in the same order
if( !all(row.names(cwm_obs_sim)==cwm_rand_mean$plot) ){
  print("WARNING - cwm_obs_sim ROWS DO NOT MATCH CWM_MEAN ROWS")
}

## Calculate SES values
cwm_ses_sim <- (cwm_obs_sim - select(cwm_rand_mean, -plot)) / select(cwm_rand_sd, -plot)

## Make a logical matrix
cwm_ses_sim_sig <- abs(cwm_ses_sim) >= 1.96
colSums(cwm_ses_sim_sig)

```

Based on these results, and the very small number of plots whose CWM values that are significant when considering the null distribution, it seems that there is little evidence for environmental filtering here.

```{r}
apply(cwm_ses_sim, MARGIN = 2, mean)
apply(cwm_ses_sim, MARGIN = 2, t.test)

```

Grouping all values together, there's some indication that FWC is gnerally less random than expected by the null model, so perhaps there is some environmental filtering for this trait going on.

Plot the distributions of the SES values for each trait

```{r}
cwm_ses_temp <- cwm_ses_sim
cwm_ses_temp$plot <- row.names(cwm_ses_sim)
cwm_ses_temp_long <- tidyr::gather(cwm_ses_temp, "trait", "value", 1:3)

ggplot(data = cwm_ses_temp_long, aes(x = value)) +
  geom_histogram() +
  geom_vline(xintercept = 0) +
  facet_wrap(~trait) +
  theme_bw()

```


# RLQ for traits, abundance, and environment


```{r}
## Correspondence analysis of the L matrix/table 
l_sim_coa <- dudi.coa(l_sim_mat, scannf = FALSE)

## PCA of the R and Q data tables
r_sim_pca <- dudi.pca(scale(r_sim_mat), row.w = l_sim_coa$lw, scannf = FALSE)
q_sim_pca <- dudi.pca(scale(q_sim_mat), row.w = l_sim_coa$cw, scannf = FALSE)

## Run RLQ function
sim_rlq <- rlq(r_sim_pca, l_sim_coa, q_sim_pca, scannf = FALSE)


## Plot the result
plot(sim_rlq)
```


Look at that bi-plot of just the trait and environmental variables.

```{r}
s.arrow(sim_rlq$c1, xlim=c(-1,1), boxes = FALSE)
s.label(sim_rlq$li, add.plot=T, clab=1.5)
```


# Fourth-Corner

```{r}
nrepet <- 999 
sim_fourthcorner <- fourthcorner(as.data.frame(scale(r_sim_mat)), 
                                 l_sim_mat, 
                                 as.data.frame(scale(q_sim_mat)), 
                                 modeltype = 6,
                                 p.adjust.method.G = "none", 
                                 p.adjust.method.D = "none", 
                                 nrepet = nrepet)

summary(sim_fourthcorner)

plot(sim_fourthcorner, alpha = 0.05, col = c("white", "grey", "black"))

```



