---
title: "Trait Analysis Example"
output: 
  html_notebook: 
    toc: yes
---

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.


In this document, we show several trait modeling analyses using data collected in the Baviaanksloof region of the Cape Floristic Region in South Africa.

```{r, message=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(GGally)
library(ade4)
library(vegan)
library(reshape2)
library(FD)
```

Read in the trait data sets.
First need to read in full CFR dataset, then filter to B'kloof data.
Following the notation of Dray and colleagues (REFS), these two datasets are used to construct the **Q** matrix.
The **Q** matrix is the $p \times s$ table describing $s$ traits for $p$ species. 
By convention, each species is assigned a single value for each trait, usually an average if the trait is measured on a continuous scale.
In our datasets, we have multiple measurements per species, so we must *reduce* our data, resulting in a loss of some of our collected information.
Essentially, this process ignores *intra-specific* variation.
It remains unclear what role *intra-specific* variation plays in determining the patterns of trait by environment relationships in real communities.

```{r}
cfr_traits_lab <- read.csv("/Users/mlammens/Dropbox/UConn-PostDoc/Projects/Dimensions-GCFR/Dimensions-ZA/data_base/individualsXtraits_lab.csv")
cfr_traits_field <- read.csv("/Users/mlammens/Dropbox/UConn-PostDoc/Projects/Dimensions-GCFR/Dimensions-ZA/data_base/individualsXtraits_field.csv")

bk_traits_lab <- filter(cfr_traits_lab, region == "baviaanskloof")
bk_traits_field <- filter(cfr_traits_field, region == "baviaanskloof")
```

## Summary information and statistics of trait data

Number of species - lab

```{r}
length(unique(bk_traits_lab$species))
```

Number of species - field

```{r}
length(unique(bk_traits_field$species))
```

These are equal, which is a good check.

### Select a set of traits to investigate

For the examples in this chapter, we will use the same set of traits used in Aiello-Lammens et al. 2016.

#### Field measurements

* Plant height
* Canopy area
* Branch order
* Resprout post-fire

#### Lab measurements

* Leaf-mass area (LMA)
* Leaf length-width ratio (LWR)
* Leaf lamina thickness
* Maximum leaf lamina width
* Leaf freshwater content (FWC)
* Stem freshwater content (twFWC)

```{r}
obs_ids <- c("uid", "species", "latitude", "longitude", "region")

lab_traits <- c("lma", "lwr", "leaf_thickness_mm", "max_leaf_width_cm",
                "fwc", "twig_fwc")

field_traits <- c("height_cm", "canopy_area_cm2")
```


### Create complete cases data sets

```{r}
bk_traits_lab_sub <- 
  bk_traits_lab %>%
  select(which(names(bk_traits_lab) %in% c(obs_ids, lab_traits)))

bk_traits_lab_complete <- 
  bk_traits_lab_sub[complete.cases(bk_traits_lab_sub[lab_traits]), ]

bk_traits_field_sub <- 
  bk_traits_field %>%
  select(which(names(bk_traits_field) %in% c(obs_ids, field_traits)))

bk_traits_field_complete <- 
  bk_traits_field_sub[complete.cases(bk_traits_field_sub[field_traits]), ]

```


# Data visualization

First look at data in a non-aggregated way. 
Each data point represents measurements taken from samples collected from individual plants.

## Trait-by-trait plots

Using the `GGally::ggpairs` function.

### Lab traits

```{r}
ggpairs(bk_traits_lab_complete[lab_traits])
```

Based on this data visualization, we see that we should perform data transformation.
As is common in plant ecology, we apply a log~10~ transfromation.

```{r}
bk_traits_lab_complete_log10 <- bk_traits_lab_complete
bk_traits_lab_complete_log10[lab_traits] <- log10(bk_traits_lab_complete_log10[lab_traits])
```

After log~10~ transformaion, plot trait-trait relationships.

```{r}
ggpairs(bk_traits_lab_complete_log10[lab_traits])
# figon ggsave("../manuscript/figures/cfr_lab_traits.eps", device = "eps")
```

### Field traits

Look at trait-trait relationships for field traits.
The field traits differ from the lab traits in that field traits were measured only on one "representative" individual for each species observed at a data collection site.


```{r}
ggpairs(bk_traits_field_complete[field_traits])
```

As with the lab data, here we should trasform the measurements of `height_cm` and `canopy_area_cm2`.

```{r}
bk_traits_field_complete_log10 <- bk_traits_field_complete
bk_traits_field_complete_log10[c("height_cm", "canopy_area_cm2")] <- log10(bk_traits_field_complete[c("height_cm", "canopy_area_cm2")])
```

Replot.

```{r}
ggpairs(bk_traits_field_complete_log10[field_traits])
```


# PCA of individual-level traits

## Lab traits

In order to reduce dimensionality of the lab trait data set, we can apply principal component analysis. 

```{r}
bk_traits_lab_pca <- rda(bk_traits_lab_complete_log10[lab_traits], scale = TRUE)
```

Plot biplot and examine summary.

```{r}
bk_traits_lab_pca
summary(bk_traits_lab_pca, scaling = 2)
summary(bk_traits_lab_pca, scaling = 1)
```

```{r}
biplot(bk_traits_lab_pca, scaling = 2, main = "PCA - scaling = 2", type = c("text", "points"))
biplot(bk_traits_lab_pca, scaling = 2, choices = c(1,3), main = "PCA - scaling = 2", type = c("text", "points"))
```

```{r}
biplot(bk_traits_lab_pca, scaling = 1, main = "PCA - scaling = 1", type = c("text", "points"))
```

## Field traits

Given the already small number of traits collected in the field, and the fact that only two of the three are continuous, dimension reduction is unnecessary.

# Species-level trait values

As outlined in the main text of the chapter, there are many challenges and problems presented when *intraspecific* trait variation is either ignored or not considered.
However, most of the 'standard' trait analyses (e.g., Fourth-corner problem, RLQ, etc.) use only species-level trait values.
Furthemore, to some degree, the assumption of indpendence among samples is violated when we combine both intra- and inter-specific sources of variation, as we have done in the PCA of lab traits, shown above. 
For these reasons, it's worth calculating species-level mean trait values.

I will use the log-transformed values and complete data sets used in the PCA above in calculating species-level mean trait values.

## Species-level lab traits

```{r}
bk_traits_lab_species <-
  bk_traits_lab_complete_log10 %>%
  group_by(species) %>%
  dplyr::summarise(max_leaf_width_cm = mean(max_leaf_width_cm),
                   leaf_thickness_mm = mean(leaf_thickness_mm),
                   lma = mean(lma),
                   fwc = mean(fwc),
                   lwr = mean(lwr),
                   twig_fwc = mean(twig_fwc))

ggpairs(select(bk_traits_lab_species, -species))
```

## Species-level field traits

Note that in this case, I am treating branch order *not* as an ordered variable, but as a numeric value. This results in a continuous **branch order** value, which while not logical for a specific plant, does make sense when considering the *average* branch order value for a species (or a plot, as used in CWM).
Unfortunately, this conversion yields a bimodal distribution, since most plants were observed with a branch order of either 0, 1, or 4.

```{r}
bk_traits_field_species <-
  bk_traits_field_complete_log10 %>%
  group_by(species) %>%
  dplyr::summarise(height_cm = mean(height_cm),
                   canopy_area_cm2 = mean(canopy_area_cm2))

ggpairs(select(bk_traits_field_species, -species))
```

## Combine lab and field trait data

Now we can combine these two datasets.

```{r}
bk_traits_species <- 
  dplyr::inner_join(bk_traits_field_species, bk_traits_lab_species, by = "species")

summary(bk_traits_species)
```

Only considering those species for which we have **complete** data, our combined trait data set contains **nine trait measurements** for `r nrow(bk_traits_species)` species.

# PCA of species-level trait data

We can now perform a PCA of the species-level trait data.
Note that since we have calculated species-level trait mean values, we are no longer confounding intra- and inter-specific trait variation in this anlaysis.

```{r}
bk_traits_species_pca <- rda(bk_traits_species[c(field_traits,lab_traits)], scale = TRUE)
```

Plot biplot and examine summary.

```{r}
bk_traits_species_pca
summary(bk_traits_species_pca, scaling = 2)
summary(bk_traits_species_pca, scaling = 1)
```


```{r}
biplot(bk_traits_species_pca, scaling = 2, main = "PCA - scaling = 2", type = c("text", "points"))
biplot(bk_traits_species_pca, scaling = 2, choices = c(1,3), main = "PCA - scaling = 2", type = c("text", "points"))
```



## Read in environmental data

We have two sets of environmental data to consider.
The first is the observations of environmental conditions at the sites were trait data were collected.
The second is the observations of environmental conditions at the *plots* were species relative percent cover values were collected. 
These locations were nearby to each other, but not exactly matching.
The dataset of environmental conditions at the plots corresponds to the **R** matrix, following the notation of Dray and colleagues. 
This matrix is an $n \times m$ table of measurements of $m$ environmental variables taken at $n$ plots or sites.


```{r}
cfr_env_plots <- read.csv("/Users/mlammens/Dropbox/UConn-PostDoc/Projects/Dimensions-GCFR/Dimensions-ZA/data_base/plotXenv.csv")
cfr_env_fieldsites <- read.csv("/Users/mlammens/Dropbox/UConn-PostDoc/Projects/Dimensions-GCFR/Dimensions-ZA/data_base/field_collXenv.csv")
bk_env_plots <- filter(cfr_env_plots, region == "baviaanskloof")
bk_env_fieldsites <- filter(cfr_env_fieldsites, region == "baviaanskloof")
```

### Summary information and variable selection

For this example analysis, we're going to rely on the environmental variable selection already used in Aiello-Lammens et al. (2016). 
The environmental variables used included minimum temperature in July (MTmin.07), maximum temperature in January (MTmax.01), aspect (separately as northerness and easterness), solar radiation, and elevation.

```{r}
plot_ids <- c("plot_smpl", "latitutde", "longitude", "region")
fieldsite_ids <- c("latitude", "longitude", "region")
env_vars <- c("MTmin.07", "MTmax.01", "solar_rad", "altitude")
```

#### Create complete cases data sets

```{r}
bk_env_plots_sub <- 
  bk_env_plots %>%
  select(which(names(bk_env_plots) %in% c(plot_ids, env_vars)))
bk_env_plots_complete <-
  bk_env_plots_sub[complete.cases(bk_env_plots_sub), ]

bk_env_fieldsites_sub <- 
  bk_env_fieldsites %>%
  select(which(names(bk_env_fieldsites) %in% c(fieldsite_ids, env_vars)))
bk_env_fieldsites_complete <-
  bk_env_fieldsites_sub[complete.cases(bk_env_fieldsites_sub), ]

```


#### EDA of environmental data

Use `ggpairs` to examine distributional patterns of environmental variables and correlation among variables.

##### Environmental variables at Plots

```{r}
ggpairs(bk_env_plots_complete[env_vars])
```


##### Environmental variables at Field Sites

Note that there is a 'circular' relationship between easterness and northerness.
This is because they are derived from the same environmental value (aspect). 
We also must note that there are strong correlations between MTmin.07 and MTmax.01, and between both of these variables and Elevation.
This is to be expected, and basically shows that as elevation goes up, temperature goes down, and locations with warmer (colder) winters have warmer (colder) summers.

```{r}
ggpairs(bk_env_fieldsites_complete[env_vars])
```

# Individual-level Trait by environment relationships - field sites

We can look for associations between trait values and environmental values at the field sites where trait measurments were made. 
This form of analysis is important in trait by environment analyses, but generally not available if species-level trait values are being used.
That is, this kind of analysis is best done for individuals.

### Make combined trait and field site environment datasets

```{r}
bk_traits_field_env <- 
  inner_join(bk_traits_field_complete_log10, bk_env_fieldsites_complete)
bk_traits_lab_env <-
  inner_join(bk_traits_lab_complete_log10, bk_env_fieldsites_complete)
```

### Make simple plots of trait x environment relationships

#### Field traits

```{r}
bk_traits_field_env_m <-
  melt(bk_traits_field_env, measure.vars = c("height_cm", "canopy_area_cm2"), variable.name = "Trait", value.name = "Trait.Val")
bk_traits_field_env_m <-
  melt(bk_traits_field_env_m, measure.vars = env_vars, variable.name = "Env", value.name = "Env.Val")
```

Make plot

```{r}
ggplot(bk_traits_field_env_m, aes(x = Env.Val, y = Trait.Val)) +
  geom_point() +
  geom_smooth() +
  facet_grid(Trait ~ Env, scales = "free")
```

#### Lab traits

```{r}
bk_traits_lab_env_m <-
  melt(bk_traits_lab_env, measure.vars = lab_traits, variable.name = "Trait", value.name = "Trait.Val")
bk_traits_lab_env_m <-
  melt(bk_traits_lab_env_m, measure.vars = env_vars, variable.name = "Env", value.name = "Env.Val")
```

Make plot

```{r}
ggplot(bk_traits_lab_env_m, aes(x = Env.Val, y = Trait.Val)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(Trait ~ Env, scales = "free")
```

### Linear regression models

```{r}
traitXenv_ols_models <- 
  bk_traits_lab_env_m %>%
  group_by(Trait, Env ) %>% 
  do( mod = lm( Trait.Val ~ Env.Val, data = . ) )

traitXenv_ols_models <- 
  mutate( traitXenv_ols_models, r_sqrd = summary(mod)$adj.r.squared )

traitXenv_ols_models_summary <-
  traitXenv_ols_models %>% 
  do(data.frame( var = names(coef(.$mod)),
                 coef(summary(.$mod))))

```

# Species-level trait by environment variables

This analysis applies only to the lab trait data. At each site, multiple individuals for each species were sampled to measure leaf traits. By grouping at the field-site and species levels, we can ignore intra-species - intra-site level variation.

```{r}
bk_traits_lab_species_fieldsite <-
  bk_traits_lab_env %>%
  group_by(uid, species) %>%
  summarise(latitude = unique(latitude), longitude = unique(longitude), region = unique(region),
            max_leaf_width_cm = mean(max_leaf_width_cm),
            leaf_thickness_mm = mean(leaf_thickness_mm),
            lma = mean(lma),
            fwc = mean(fwc),
            lwr = mean(lwr),
            twig_fwc = mean(twig_fwc),
            MTmax.01 = unique(MTmax.01),
            MTmin.07 = unique(MTmin.07),
            altitude = unique(altitude),
            solar_rad = unique(solar_rad))
```

Melt dataset

```{r}
bk_traits_lab_species_fieldsite_m <-
  melt(bk_traits_lab_species_fieldsite, measure.vars = lab_traits, variable.name = "Trait", value.name = "Trait.Val")
bk_traits_lab_species_fieldsite_m <-
  melt(bk_traits_lab_species_fieldsite_m, measure.vars = env_vars, variable.name = "Env", value.name = "Env.Val")
```

Make plot

```{r}
ggplot(bk_traits_lab_species_fieldsite_m, aes(x = Env.Val, y = Trait.Val)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(Trait ~ Env, scales = "free")
```


# Community Level Data

## Read in the community level data

The community level data are from surveys that measured relative percent cover for each species in 10 x 5 meter plots in the Baviaanskloof region. 
In this case, we will only use the "A" plots.


```{r}
plotXspecies_cover <-
  read.csv("/Users/mlammens/Dropbox/UConn-PostDoc/Projects/Dimensions-GCFR/Dimensions-ZA/data_base/plotXspecies_cover.csv")
bk_plotXspecies_cover <- filter(plotXspecies_cover, region == "baviaanskloof", historic == 0)
bk_plotXspecies_cover <- bk_plotXspecies_cover[ which(grepl(pattern = "a", x = bk_plotXspecies_cover$plot)), ]
bk_plotXspecies_cover <- 
  bk_plotXspecies_cover %>%
  select(plot_smpl, latitude, longitude, species, perc_cover, rel_perc_cover)
```

In order to convert this into the **L** matrix, following the notation of Dray et al., we need to reshape the dataset from long to wide. 
The **L** matrix is a $n \times p$ table of abundances (or occurrences) of $p$ species in $n$ sites (or plots).

```{r}
pXs <- dcast(data = bk_plotXspecies_cover, 
             formula = plot_smpl ~ species,
             value.var = "rel_perc_cover")

## Convert nas to 0
pXs[ is.na(pXs) ] <- 0

## Assign plots to row.names
row.names( pXs ) <- pXs$plot_smpl
pXs <- dplyr::select( pXs, -plot_smpl )

```


### Determine which species are in both the trait and plot datasets

```{r}
spec_both <- intersect( bk_traits_species$species, names(pXs) )

## Extract only those species with traits from both the pXs and tXs data.frames
pXs_forCWM <- pXs[ spec_both ]
tXs_forCWM <- filter(bk_traits_species, species %in% spec_both )

# ## Remove plots that are "empty" after removing species w/o traits
# pXs_forCWM <- pXs_forCWM[ -which( apply( pXs_forCWM, MARGIN = 1, FUN = sum ) == 0 ), ]

## Name rows of tXs_forCWM data.frame
row.names( tXs_forCWM ) <- tXs_forCWM$species
tXs_forCWM$species <- NULL

```



# RDA of traits and environment


# CWM of traits

Use the `FD` package to calculate Community Weighted Mean trait values.

```{r}
bk_dbFD_traits <-
  dbFD( x = tXs_forCWM, 
        a = pXs_forCWM,
        w.abun = TRUE, 
        calc.FDiv = TRUE, 
        calc.CWM = TRUE,
        corr = "cailliez" )

```


```{r}
## Get the CWM value data.frame
bk_cwm_traits <- bk_dbFD_traits$CWM
bk_cwm_traits$plot <- row.names( bk_cwm_traits )

## Use only complete cases
bk_cwm_traits <- bk_cwm_traits[ complete.cases(bk_cwm_traits), ]

```



### PCA of CWM trait values

```{r}
## Calculate PCA of CWM values
bk_cwm_traits_pca <- rda( bk_cwm_traits[ c(lab_traits, field_traits) ], scale = TRUE )
bk_cwm_traits_pca$CA$eig / sum( bk_cwm_traits_pca$CA$eig )
cumsum( bk_cwm_traits_pca$CA$eig / sum( bk_cwm_traits_pca$CA$eig ) )

## Extract various PCA values and scores for plotting in ggplot
## -------------------------------------------------------------------------- ##

## Extract the PCA scores
bk_cwm_traits_pca_scores <- scores(bk_cwm_traits_pca, choices = 1:length(c(lab_traits, field_traits)))$sites

## Convert to data.frame
bk_cwm_traits_pca_scores <- as.data.frame( bk_cwm_traits_pca_scores )
bk_cwm_traits_vars_pca <- names( bk_cwm_traits_pca_scores )

## Add the pca scores back to the bk_cwm_traits data.frame
bk_cwm_traits <- cbind( bk_cwm_traits, bk_cwm_traits_pca_scores )

## Extract loadings
bk_cwm_traits_pca_loadings <- 
  scores(bk_cwm_traits_pca, choices = 1:length(c(lab_traits, field_traits)))$species 

## Convert to data.frame
bk_cwm_traits_pca_loadings <- as.data.frame( bk_cwm_traits_pca_loadings )
## Add a column for env_var
bk_cwm_traits_pca_loadings$env_var <- row.names(bk_cwm_traits_pca_loadings)


```

```{r}
ggplot() +
  geom_segment( data = bk_cwm_traits_pca_loadings, 
                aes( x = 0, xend = PC1*.2,
                     y = 0, yend = PC2*.2 ),
                arrow = arrow( length = unit(0.25, "cm" )), 
                colour = "black") +      
  geom_point( data = bk_cwm_traits, 
              aes( x = PC1, y = PC2 ),
              alpha = 0.6 ) +
 geom_text( data = bk_cwm_traits_pca_loadings, 
             aes(x= PC1*.2, y = PC2*.2,
                 label = env_var ),
             size = 4,
             hjust = 0.5 ) + 
  coord_equal() +
  xlab( "PC1" ) +
  ylab( "PC2" ) +
  theme_bw() +
  theme( text = element_text(family = "Times", size = 18) )
```




# RLQ for traits, abundance, and environment


```{r}
names(cfr_traits_field)
summary(cfr_traits_lab)
```



