---
title: "Trait Analysis CFR Example"
output:
  html_document:
    toc: yes
  html_notebook:
    toc: yes
---

This document accompanies the book chapter **Statistical Methods for Modeling Traits**, by Matthew Aiello-Lammens and John Silander, appearing in the **Handbook of Environmental and Ecological Statistics**, edited by Gelfand, Fuentes, Hoeting, and Smith.

Here we show the trait modeling analyses discussed in the corresponding chapter, as applied to data collected in the Baviaanksloof region of the Cape Floristic Region in South Africa.

```{r, message=FALSE}
library(ggplot2)
library(plyr)
library(dplyr)
library(GGally)
library(ade4)
library(vegan)
library(reshape2)
library(FD)
library(picante)

knitr::opts_chunk$set(fig.width=6, fig.height=6) 
```

# Read in CFR trait data

Begin by reading in the CFR individual trait data set.

```{r}
indXtraits_cfr <- read.csv("individualsXtraits_cfr.csv")
```

Determine the number of species in the CFR data set.

```{r}
length(unique(indXtraits_cfr$species))
```

## Select a set of traits to investigate

For the examples in this chapter, we will use a subset of the traits used in Aiello-Lammens et al. 2016.

* Leaf-mass area (LMA)
* Leaf length-width ratio (LWR)
* Leaf lamina thickness
* Maximum leaf lamina width
* Leaf freshwater content (FWC)
* Stem freshwater content (twFWC)

```{r}
obs_ids <- c("uid", "species", "latitude", "longitude", "region")

traits <- c("lma", "lwr", "leaf_thickness_mm", "max_leaf_width_cm",
            "fwc", "twig_fwc")
```


### Create complete cases data set

We will only use observations from individuals where the data for our set of traits is complete.

```{r}
indXtraits_cfr_sub <- 
  indXtraits_cfr %>%
  select(which(names(indXtraits_cfr) %in% c(obs_ids, traits)))

indXtraits_cfr_complete <- 
  indXtraits_cfr_sub[complete.cases(indXtraits_cfr_sub[traits]), ]
```

### Change names of traits 

This step is to make prettier plots.

```{r}
traits <- c("LMA", "LWR", "Lf_Thick", "Lf_Width", "FWC", "Twig_FWC")
names(indXtraits_cfr_complete) <- c("uid", "species", "latitude", "longitude", "region",
                                    "Lf_Width", "Lf_Thick", "LMA", "FWC", "LWR", "Twig_FWC")
```


# Exploratory Data Analysis

Use data visualization and dimension reduction to investigate large patterns in the trait data.
In this case, each data point represents trait measurements taken on individual plants.

## Trait-by-trait plots

Using the `GGally::ggpairs` function.

```{r}
ggpairs(indXtraits_cfr_complete[traits])
```

Based on this data visualization, we see that we should perform data transformation.
As is common in plant ecology, we apply a log~10~ transfromation.

```{r}
indXtraits_cfr_complete_log10 <- indXtraits_cfr_complete
indXtraits_cfr_complete_log10[traits] <- log10(indXtraits_cfr_complete_log10[traits])
```

After log~10~ transformaion, plot trait-trait relationships.

```{r}
ggpairs(indXtraits_cfr_complete_log10[traits])
#ggsave("../manuscript/figures/cfr_traits.png", device = "png", width = 5, height = 5, units = "in")
```


## PCA of individual-level traits

In order to reduce dimensionality of the lab trait data set, we can apply principal component analysis. 
We carry out a PCA of the simulation data set using the `rda` function from the `vegan` package.
Note that the option `scale = TRUE` is chosen here, resulting in the data set being scaled to unit variance.

```{r}
indXtraits_cfr_pca <- rda(indXtraits_cfr_complete_log10[traits], scale = TRUE)
```

Examine summary.

```{r}
indXtraits_cfr_pca
```

```{r}
summary(indXtraits_cfr_pca, scaling = 1)
```

Examine biplots using `scaling = 1` and `scaling = 2`

### Individual-level PCA biplot - scaling = 1

```{r}
biplot(indXtraits_cfr_pca, scaling = 1, main = "PCA - scaling = 1", type = c("text", "points"), col = "black")
```

### Individual-level PCA biplot - scaling = 2

```{r}
biplot(indXtraits_cfr_pca, scaling = 2, main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
biplot(indXtraits_cfr_pca, scaling = 2, choices = c(1,3), main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
```


# Species-level trait values

As outlined in the main text of the chapter, there are many challenges and problems presented when *intraspecific* trait variation is either ignored or not considered.
However, most of the 'standard' trait analyses (e.g., Fourth-corner problem, RLQ, etc.) use only **species-level** trait values.
For these reasons, it's worth calculating species-level mean trait values, i.e. the Q matrix described above.
The **Q** matrix is the $p \times s$ table describing $s$ traits for $p$ species. 
By convention, each species is assigned a single value for each trait, usually an average if the trait is measured on a continuous scale.
In our dataset, we have multiple measurements per species, so we must *reduce* our data, resulting in a loss of some of our collected information.

## Create a **Q** matrix

```{r}
q_cfr <-
  indXtraits_cfr_complete_log10 %>%
  group_by(species) %>%
  dplyr::summarise(LMA = mean(LMA),
                   LWR = mean(LWR),
                   Lf_Thick = mean(Lf_Thick),
                   Lf_Width = mean(Lf_Width),
                   FWC = mean(FWC),
                   Twig_FWC = mean(Twig_FWC))
```


### Species-level trait-trait relationships

```{r}
ggpairs(select(q_cfr, -species))
```


### PCA of species-level trait data

We can now perform a PCA of the species-level trait data.
Note that since we have calculated species-level trait mean values, we are no longer confounding intra- and inter-specific trait variation in this anlaysis.

```{r}
specXtraits_cfr_pca <- rda(q_cfr[traits], scale = TRUE)
```

Examine summary.

```{r}
specXtraits_cfr_pca
```

```{r}
summary(specXtraits_cfr_pca, scaling = 1)
```

Examine biplots using `scaling = 1` and `scaling = 2`

### Species-level PCA biplot - scaling = 1

```{r}
biplot(specXtraits_cfr_pca, scaling = 1, main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
```

### Species-level PCA biplot - scaling = 2

```{r}
biplot(specXtraits_cfr_pca, scaling = 2, main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
biplot(specXtraits_cfr_pca, scaling = 2, choices = c(1,3), main = "PCA - scaling = 2", type = c("text", "points"), col = "black")
```

***

# Read in environmental data

We have two sets of environmental data to consider for the CFR data set.
The first is the observations of environmental conditions at the sites were trait data were collected.
The second is the observations of environmental conditions at the *plots* were species relative percent cover values were collected. 
These locations were nearby to each other, but not exactly matching.
The dataset of environmental conditions at the plots corresponds to the **R** matrix, an $n \times m$ table of measurements of $m$ environmental variables taken at $n$ plots or sites.

```{r}
plotXenv_cfr_complete <- read.csv("plotXenv_cfr.csv")
fieldsiteXenv_cfr_complete <- read.csv("fieldsiteXenv_cfr.csv")
```

## Summary information and variable selection

For this example analysis, we're going to rely on the environmental variable selection already used in Aiello-Lammens et al. (2016). 
The environmental variables used included minimum temperature in July (MTmin.07), maximum temperature in January (MTmax.01), aspect (separately as northerness and easterness), solar radiation, and elevation.

```{r}
plot_ids <- c("plot_smpl", "latitude", "longitude", "region")
fieldsite_ids <- c("latitude", "longitude", "region")
env_vars <- c("MTmin.07", "MTmax.01", "solar_rad", "altitude")
```

### Standardize (Center and Scale) the environmental variables

```{r}
plotXenv_cfr_complete[env_vars] <- scale(plotXenv_cfr_complete[env_vars])
fieldsiteXenv_cfr_complete[env_vars] <- scale(fieldsiteXenv_cfr_complete[env_vars])
```

### Change names for prettier plots

```{r}
env_vars <- c("MTmin.07", "MTmax.01", "Solar_Rad", "Elevation")
plotXenv_cfr_complete <- dplyr::rename(plotXenv_cfr_complete, Solar_Rad = solar_rad, Elevation = altitude)
fieldsiteXenv_cfr_complete <- dplyr::rename(fieldsiteXenv_cfr_complete, Solar_Rad = solar_rad, Elevation = altitude)
```

### Set R CFR matrix

```{r}
r_cfr <- dplyr::select(plotXenv_cfr_complete, -latitude, -longitude, -region)
```


## Exploratory data analysis of environmental data

Use `ggpairs` to examine distributional patterns of environmental variables and correlation among variables.

#### Environmental variables at Plots

```{r}
ggpairs(r_cfr[env_vars])
```


#### Environmental variables at Field Sites

Note that there are strong correlations between MTmin.07 and MTmax.01, and between both of these variables and Elevation.
This is to be expected, and basically shows that as elevation goes up, temperature goes down, and locations with warmer (colder) winters have warmer (colder) summers.

```{r}
ggpairs(fieldsiteXenv_cfr_complete[env_vars])
```

# Individual-level Trait by environment relationships 

We can look for associations between trait values and environmental values at the field sites where trait measurments were made. 
This form of analysis is important in trait by environment analyses, but generally not available if species-level trait values are being used.

## Make combined trait and field site environment datasets

```{r}
indXtraits_cfr_env <-
  inner_join(indXtraits_cfr_complete_log10, fieldsiteXenv_cfr_complete)
```

```{r}
indXtraits_cfr_env_m <-
  melt(indXtraits_cfr_env, measure.vars = traits, variable.name = "Trait", value.name = "Trait.Val")
indXtraits_cfr_env_m <-
  melt(indXtraits_cfr_env_m, measure.vars = env_vars, variable.name = "Env", value.name = "Env.Val")
```

### Plot individual-level trait-by-environment relationships

```{r}
ggplot(indXtraits_cfr_env_m, aes(x = Env.Val, y = Trait.Val)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_grid(Trait ~ Env, scales = "free")
```

### Linear regression models

Run one-at-a-time trait-by-environment OLS models

```{r}
traitXenv_ols_models_cfr <- 
  indXtraits_cfr_env_m %>%
  group_by(Trait, Env ) %>% 
  do( mod = lm( Trait.Val ~ Env.Val, data = . ) )

traitXenv_ols_models_cfr <- 
  mutate( traitXenv_ols_models_cfr, r_sqrd = summary(mod)$adj.r.squared )

traitXenv_ols_models_summary_cfr <-
  traitXenv_ols_models_cfr %>% 
  do(data.frame( var = names(coef(.$mod)),
                 coef(summary(.$mod))))

```

```{r}
print(traitXenv_ols_models_summary_cfr)
```


## Individual-level RDA

As outlined in the chapter, we can apply RDA to the individual-level trait-environment data.

Run RDA using formula notation.
Note that here we set `scale = TRUE` to scale the response data.

```{r}
bk_traits_ind_rda <- 
  rda(formula = indXtraits_cfr_env[traits] ~ MTmax.01 + MTmin.07 + Elevation + Solar_Rad, 
      data = indXtraits_cfr_env, scale = TRUE)
```

Examine summary.

```{r}
summary(bk_traits_ind_rda)
```

Tri-plot - scaling = 1

```{r}
plot( bk_traits_ind_rda, type='none', scaling = 1 )
points( bk_traits_ind_rda, pch=20, cex=0.8, scaling = 1 )
text(bk_traits_ind_rda,display='species',col='black',axis.bp=TRUE, scaling = 1)
text(bk_traits_ind_rda,display='bp', scaling = 1)
```

Tri-plot - scaling = 2

```{r}
plot( bk_traits_ind_rda, type='none', scaling = 2 )
points( bk_traits_ind_rda, pch=20, cex=0.8, scaling = 2 )
text(bk_traits_ind_rda,display='species',col='black',axis.bp=TRUE, scaling = 2)
text(bk_traits_ind_rda,display='bp', scaling = 2)
```

Get coefficients for RDA

```{r}
coef(bk_traits_ind_rda)
```


Get R^2 for RDA

```{r}
(R2adj <- RsquareAdj(bk_traits_ind_rda))
```

Note that the R-squared value for these traits based on these environmental variables is less than 5%.

Perform permutation test on this RDA

```{r}
anova(bk_traits_ind_rda)
```


***

# Community Level Data

## Read in the plot-level abundance data

This data set is equivelant to the **L** matrix, which is an $n \times p$ table of abundances (or occurrences) of $p$ species in $n$ sites (or plots).
For the CFR data set the community level data are from surveys that measured relative percent cover for each species in 10 x 5 meter plots in the Baviaanskloof region. 


```{r}
plotXspecies_cover_cfr <-  read.csv("plotXspec_abund_cfr.csv")
```

In order to convert this into the **L** matrix, we need to reshape the dataset from long to wide. 

```{r}
l_cfr <- dcast(data = plotXspecies_cover_cfr, 
               formula = plot_smpl ~ species,
               value.var = "rel_perc_cover")

## Convert nas to 0
l_cfr[ is.na(l_cfr) ] <- 0

## Assign plots to row.names
row.names( l_cfr ) <- l_cfr$plot_smpl
l_cfr <- dplyr::select( l_cfr, -plot_smpl )
```


### Determine which species are in both the trait and plot datasets

```{r}
spec_both <- intersect( q_cfr$species, names(l_cfr) )

## Extract only those species with traits from both the Q and L data.frames
l_cfr <- l_cfr[ spec_both ]
q_cfr_sub <- filter(q_cfr, species %in% spec_both )

## Name rows of q_cfr_sub data.frame
q_cfr_sub <- as.data.frame(q_cfr_sub)
row.names( q_cfr_sub ) <- q_cfr_sub$species
q_cfr_sub$species <- NULL

```


## CWM trait values

Use the `FD` package to calculate Community Weighted Mean trait values.
Note that the row names in Q must match the column names in L.


```{r}
cfr_dbFD_traits <-
  dbFD( x = q_cfr_sub, 
        a = l_cfr,
        w.abun = TRUE, 
        calc.FDiv = TRUE, 
        calc.CWM = TRUE,
        corr = "cailliez" )

```


```{r}
## Get the CWM value data.frame
cwm_traits_cfr <- cfr_dbFD_traits$CWM
cwm_traits_cfr$plot <- row.names( cwm_traits_cfr )
```

### Plot CWM trait-trait relationships

```{r}
ggpairs(select(cwm_traits_cfr, -plot))
```


## RDA of CWM traits with environment (CWM-RDA)

Perform RDA on the calculated CWMs.

Add the plot environmental values to the CWM matrix

```{r}
cwm_traits_cfr_env <- inner_join(cwm_traits_cfr, r_cfr, by = c("plot" = "plot_smpl"))
```


```{r}
cwm_traits_cfr_rda <- 
  rda(formula = cwm_traits_cfr[traits] ~ MTmax.01 + MTmin.07 + Elevation + Solar_Rad, 
      data = cwm_traits_cfr_env, scale = TRUE)
print(cwm_traits_cfr_rda)
```


Get coefficients and R-squared values

```{r}
coef(cwm_traits_cfr_rda)
```

```{r}
RsquareAdj(cwm_traits_cfr_rda)
```

Interestingly, not much more than we got with individuals.

### Tri-plot - scaling = 1

```{r}
plot( cwm_traits_cfr_rda, type='none', scaling = 1 )
points( cwm_traits_cfr_rda, pch=20, cex=0.8, scaling = 1 )
text(cwm_traits_cfr_rda,display='species',col='black',axis.bp=TRUE, scaling = 1)
text(cwm_traits_cfr_rda,display='bp', scaling = 1)
```

### Tri-plot - scaling = 2

```{r}
plot( cwm_traits_cfr_rda, type='none', scaling = 2 )
points( cwm_traits_cfr_rda, pch=20, cex=0.8, scaling = 2 )
text(cwm_traits_cfr_rda,display='species',col='black',axis.bp=TRUE, scaling = 2)
text(cwm_traits_cfr_rda,display='bp', scaling = 2)
```


### Modify R, L, and Q matrices for further analyses

The R, L, and Q matrices need to be modified to that they are in the exact format needed for fourth-corner and RLQ analysis.

```{r}
r_cfr_mat <- r_cfr[env_vars]
row.names(r_cfr_mat) <- r_cfr$plot

l_cfr_mat <- l_cfr
names(l_cfr_mat) <- sub(" ", ".", names(l_cfr_mat))

q_cfr_mat <- q_cfr_sub
q_cfr_mat <- as.data.frame(q_cfr_mat)
row.names(q_cfr_mat) <- names(l_cfr_mat)
```


## Randomization approaches to test if CWMs show signes of environmental filtering

In this example, we will use 'indpendent swap' ranomizaiton.
The work flow here is to apply the randomization/ permutation on the L matrix.
Then recalculate CWMs.
Store the CWM values for comparison to the original values later.

```{r, message=FALSE, eval = FALSE}
## Choose number of iterations
iter_n <- 999
cwm_cfr_rand_all <- vector()

for (iter in 1:iter_n){
  ## Shuffle the L matrix
  l_rand <- randomizeMatrix(l_cfr_mat, null.model = "independentswap")
  
  ## Calculate CWM values
  cwm_rand <- dbFD( x = q_cfr_mat, 
                    a = l_rand,
                    w.abun = TRUE, 
                    calc.FDiv = FALSE, 
                    calc.CWM = TRUE,
                    corr = "cailliez" )
  cwm_rand <- cwm_rand$CWM
  
  ## Add a column for the iteration number
  cwm_rand$iter <- iter
  cwm_rand$plot <- row.names(cwm_rand)
  
  ## Add this iteration to the full dataset
  cwm_cfr_rand_all <- rbind(cwm_cfr_rand_all, cwm_rand)
}

save(cwm_cfr_rand_all, file = "cwm_cfr_rand_all.RData")

```


After generating random CWM sets, aggregate the data.

```{r}
## Load in the saved randomization results
load("cwm_cfr_rand_all.RData")

## Calculate the mean values
cwm_rand_mean <- cwm_cfr_rand_all %>%
  group_by(plot) %>%
  summarise(LMA = mean(LMA),
            LWR = mean(LWR),
            Lf_Thick = mean(Lf_Thick),
            Lf_Width = mean(Lf_Width),
            FWC = mean(FWC),
            Twig_FWC = mean(Twig_FWC))

## Calculate the SD values
cwm_rand_sd <- cwm_cfr_rand_all %>%
  group_by(plot) %>%
  summarise(LMA = sd(LMA),
            LWR = sd(LWR),
            Lf_Thick = sd(Lf_Thick),
            Lf_Width = sd(Lf_Width),
            FWC = sd(FWC),
            Twig_FWC = sd(Twig_FWC))
```

Get the observed values and calculate SES

```{r}
cwm_obs_cfr <- select(cwm_traits_cfr, LMA:Twig_FWC )

## Check that they are in the same order
if( !all(row.names(cwm_obs_cfr)==cwm_rand_mean$plot) ){
  print("WARNING - cwm_obs_cfr ROWS DO NOT MATCH CWM_MEAN ROWS")
}

## Calculate SES values
cwm_ses_cfr <- (cwm_obs_cfr - select(cwm_rand_mean, -plot)) / select(cwm_rand_sd, -plot)

## Make a logical matrix
cwm_ses_cfr_sig <- abs(cwm_ses_cfr) >= 1.96
colSums(cwm_ses_cfr_sig)

```

Based on these results, and the very small number of plots whose CWM values that are significant when considering the null distribution, it seems that there is little evidence for environmental filtering here.

```{r}
apply(cwm_ses_cfr, MARGIN = 2, mean)
apply(cwm_ses_cfr, MARGIN = 2, t.test)

```

Grouping all values together, there's some indication that FWC is gnerally less random than expected by the null model, so perhaps there is some environmental filtering for this trait going on.

Plot the distributions of the SES values for each trait

```{r}
cwm_ses_temp <- cwm_ses_cfr
cwm_ses_temp$plot <- row.names(cwm_ses_cfr)
cwm_ses_temp_long <- tidyr::gather(cwm_ses_temp, "trait", "value", 1:6)

ggplot(data = cwm_ses_temp_long, aes(x = value)) +
  geom_histogram() +
  geom_vline(xintercept = 0) +
  facet_wrap(~trait) +
  theme_bw()

```



# RLQ for traits, abundance, and environment


```{r}
## Correspondence analysis of the L matrix/table 
l_cfr_coa <- dudi.coa(l_cfr_mat, scannf = FALSE)

## PCA of the R and Q data tables
r_cfr_pca <- dudi.pca(scale(r_cfr_mat), row.w = l_cfr_coa$lw, scannf = FALSE)
q_cfr_pca <- dudi.pca(scale(q_cfr_mat), row.w = l_cfr_coa$cw, scannf = FALSE)

## Run RLQ function
cfr_rlq <- rlq(r_cfr_pca, l_cfr_coa, q_cfr_pca, scannf = FALSE)


## Plot the result
plot(cfr_rlq)
```


Make tri-plot from RLQ results

```{r}
s.arrow(cfr_rlq$c1, xlim=c(-1,1), boxes = FALSE)
s.label(cfr_rlq$li, add.plot=T, clab=1.5)
```


# Fourth-Corner

```{r}
nrepet <- 999 
cfr_fourthcorner <- fourthcorner(as.data.frame(scale(r_cfr_mat)), 
                                 l_cfr_mat, 
                                 as.data.frame(scale(q_cfr_mat)), 
                                 modeltype = 6,
                                 p.adjust.method.G = "none", 
                                 p.adjust.method.D = "none", 
                                 nrepet = nrepet)

plot(cfr_fourthcorner, alpha = 0.05, col = c("white", "grey", "black"))

```

```{r}
print(cfr_fourthcorner)
```


Here's the fourth-corner analysis treating the L cfr_matrix as presence/absence, rather than relative percent cover.

```{r}
cfr_fourthcorner_pa <- fourthcorner(r_cfr_mat, 
                                    l_cfr_mat, 
                                    q_cfr_mat, 
                                    tr01 = TRUE,
                                    modeltype = 6,
                                    p.adjust.method.G = "none", 
                                    p.adjust.method.D = "none", 
                                    nrepet = nrepet)

plot(cfr_fourthcorner_pa, alpha = 0.05, col = c("white", "grey", "black"))

```

